Perceptions and Preferences for Meritocracy in Survey Research

### 2023-05-18 paper.Rmd
submission IJS

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies. One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather inconsistent. Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items. Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization. Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose (N=2,141). Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey. Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::word_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true # for word output
fontsize: 10pt
---

```{r eval=FALSE, include=FALSE}
 # for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
 # clean #in the yml
 rmarkdown::render("paper.Rmd", output_format = "bookdown::pdf_document2")
 rmarkdown::render("paper.Rmd", output_format = "bookdown::html_document2")
 # rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
```

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H", fig.align = 'center')
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
 # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
 options(kableExtra.auto_format = FALSE)
 rm(list=ls())
 options(knitr.kable.NA = '')
 options(knitr.graphics.error = FALSE)
 Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

@@

@@

Economic inequality and income concentration have become topics of growing concern over the last years, acquiring even more salience in times of pandemic and economic crises. This has led to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014]. In this context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020]. Within this context, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].

In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962]. This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breen_class_1999; @saunders_might_1995; @yair_meritocracy_2007; @land_we_2006; @young_rise_1962]. Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of the individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@preminger_meritocracy_2020; @trump_when_2020; @hadjar_meritokratie_2008; @madeira_primes_2019]. Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon the individual status and social mobility [@sandel_tyranny_2020; @witteveen_reconsidering_2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_meritocracy_2019; @khan_privilege_2013].

Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics. For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019]. Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy. However, as we will show later, the concepts used as well as the instruments for measuring meritocracy vary extensively among studies. In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators. Such trends not only limit the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different disciplines and lines of research.

Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two main aspects in the study of meritocracy: perceptions and preferences. Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies seem to suggest. This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.

## The black-box of meritocratic beliefs {-#blackbox}

@@

One recent definition of meritocracy by @mijs_paradox_2019 is the following: "When I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9]. In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale. The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.

The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead". This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender. Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic. As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).

In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assumes effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy? The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny. Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations. To illustrate this point let us consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery. It consists of a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts. Another version of this same battery used in other surveys - sometimes together with the previous one - asking about "how important you think it _should_ be ...", and then listing the same concepts.  Therefore, the question raised here is: Which one of both is a "belief": what _is_ or what it _should_ be?

The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992. As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30). Therefore, beliefs cover almost anything related to subjective factors. To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Several papers dealing with meritocracy use the term beliefs (i.e. what should be), while actually referring to perceptions (i.e. what is). This occurs for instance in @reynolds_perceptions_2014, in which the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007]. The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs). They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "... people are rewarded for their skills?".

Is the belief in meritocracy a perception or a preference with normative meaning? In order to expand the analytical conceptual framework, we believe that both dimensions should be included as proposed by @duru-bellat_who_2012. This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e. have a high correlation), or whether they are independent aspects of the same phenomenon. As @sonhing_merit_2011 has pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p. 435). In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy than to someone perceiving low meritocracy. To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].

@@

Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life. For instance, @kunovich_systems_2007 used an items' battery listing a number of factors in relation to "How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy). They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e. reverse coded). The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy. A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].

The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above. They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic. Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did. Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items. Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further empirical assessment.

@@

Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the latent construct is measured perfectly (i.e. no error or residual variance) by the selected indicators, going as far as to propose that "... In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654]. Although some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, somewhat surprisingly they finally chose to build a sum index despite proving a multidimensional latent structure.

## Conceptual proposal for studying meritocracy {-#instrumentprop}

@@

- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.

- Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.

- _Brief_, as to be used in regular public opinion surveys. In this respect, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).

@@

```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
 knitr::include_graphics("input/images/Fig1.generalf.png")
```

The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019]. Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as "Hard work is important to get ahead in society". Preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research. Non-meritocratic elements refer to the use of resources such as personal contacts or family advantages to getting ahead in life.

@@

The analysis is organized into three studies. Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data from the inequality module of the International Social Survey Programme (ISSP). Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used data in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard. Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences. Finally, in Study 3 we perform an additional validity analysis of the meritocracy scale.

# Study 1: Measuring meritocracy with international secondary data

@@

The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues. This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56021 individuals. Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, the reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far only with a restricted set of countries).

@@

There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy measurement scheme. However, it is important to mention that the items were not originally designed as a specific measure of meritocracy. The variables are presented below organized in perceptions and preferences:

```{r table-issp, echo=FALSE}
  if (!require("pacman")) install.packages("pacman")
  # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
  pacman::p_load(knitr, kableExtra, dplyr)
  table_format = if(knitr::is_html_output()) {
    "html"
  } else if(knitr::is_latex_output()) {
    "latex"
  }

@@

# (APPENDIX) Appendix {-}

>>>>

---
title: "Perceptions and Preferences for Meritocracy in Survey Research"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
params:
  lang: en
author: 
- name: Juan Carlos Castillo
  affiliation: Department of Sociology, Universidad de Chile
  email: juancastillov@uchile.cl
  number: 1
- name: Julio Iturra
  affiliation: Bremen International Graduate School of Social Sciences, Universität Bremen
  number: 2
- name: Luis Maldonado
  affiliation: Institute of Sociology, Universidad Católica de Chile
  number: 3
- name: Francisco Meneses
  affiliation: 
  number: 1
- name: Jorge Atria
  affiliation: Department of Sociology, Universidad Diego Portales, Chile
  number: 5
abstract: |
  Economic and social inequalities are causing growing concerns and crises in contemporary societies. One factor suggested to explain the persistence of inequality is the belief in meritocracy, which justifies economic disparities based on differences in effort and talent. However, empirical research on meritocracy is relatively new and characterized by diverse conceptualizations and measures, leading to inconsistent findings and interpretations. To address this issue, the present paper proposes a conceptual and measurement framework for perceptions and preferences regarding meritocracy and non-meritocracy, focusing on comparative survey research methodologies. The proposal is first tested through confirmatory factor analysis (CFA) using data from the International Social Survey Programme (ISSP) inequality module, commonly used in comparative research on meritocracy. Based on these findings,  a second study focuses on developing a novel scale that better captures the distinction between perceptions and preferences related to meritocracy. This scale is tested using a sample of 2,141 individuals from Chile. The results of this study contribute to refining the measurement of meritocracy and non-meritocracy and shed light on the importance of considering different dimensions when studying meritocracy in survey research, both conceptually and empirically.
  \newline
  \newline

   **Keywords**: meritocracy, survey methodology, social inequality, construct measurement 
output:
  bookdown::word_document2:
    number_sections: false
    reference_docx: input/template.docx
    toc: false
  bookdown::pdf_document2:
    template: null
    number_sections: false    
    toc: false
    keep_tex: false
    pandoc_args:
      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion    
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=1in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage[british,UKenglish,USenglish,american]{babel}
always_allow_html: true # for word output
fontsize: 10pt
papersize: a4
---

```{r eval=FALSE, include=FALSE}
 # for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
 # clean #in the yml
 rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::pdf_document2")
 rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::html_document2")
 rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::word_document2")
```

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = TRUE,out.width = '85%',fig.pos= "H", fig.align = 'center')
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
 # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
 options(kableExtra.auto_format = FALSE)
 rm(list=ls())
 options(knitr.kable.NA = '')
 options(knitr.graphics.error = FALSE)
 Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

<!---
\pagebreak

@@

--->

\selectlanguage{USenglish}

@@

Economic inequality and income concentration have become topics of growing concern over the last years, leading to a series of social upheavals in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].  In such a context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020]. Within this research area, and particularly in times of economic and health crises, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].

In general, meritocracy has been defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962]. This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breen_class_1999; @saunders_might_1995; @yair_meritocracy_2007; @land_we_2006; @young_rise_1962]. Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of the individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@preminger_meritocracy_2020; @trump_when_2020; @hadjar_meritokratie_2008; @madeira_primes_2019]. Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon the individual status and social mobility [@sandel_tyranny_2020; @witteveen_reconsidering_2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_meritocracy_2019; @khan_privilege_2013].

Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and (or) contextual characteristics. For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019]. Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy. However, as we will show later, the concepts and instruments used for measuring meritocracy vary extensively among studies. In many cases, similar phenomena are associated with different indicators, and conversely, different phenomena are measured with similar indicators. Such approaches limit the comparability of studies and the capacity to understand the effects of meritocratic beliefs across different disciplines and lines of research.

Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences. Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they would not be two poles of the same continuum as previous studies suggest. This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, suitable for public opinion surveys and thus allowing to integrate meritocratic beliefs in studying different social phenomena.

## The black box of meritocratic beliefs {-#blackbox}

@@

One recent definition of meritocracy by @mijs_paradox_2019 is the following: "When I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9]. In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored on a likert scale. The assumptions behind such a definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.

The item used by @mijs_paradox_2019 is part of an item's battery which is available in several international surveys, usually called "reasons to get ahead." This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender. Therefore, considering only one of the items means that other aspects associated with talent, such as education, would not be deemed meritocratic. As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).

In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only (or main) dimension behind the concept of meritocracy? We argue that whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view, but it also requires empirical evidence in studies focused on the subjective aspects of inequality. Furthermore, considering talent as part of meritocracy certainly opens some interesting avenues of research, for instance, some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020]. Therefore, we propose that a concept (and measure) of meritocracy benefits from including both talent and effort as in the original Young's conception.

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations. To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above. Some versions of this battery ask "how important you think it is for getting ahead in life" and then list several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.  Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?

The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992. As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30). Therefore, beliefs use to cover almost anything related to subjective factors. In this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e., thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e., thoughts about what should be)"(p.359). Several papers dealing with meritocracy use the term beliefs (i.e. what should be), while referring to perceptions (i.e., what is). For instance,  in @reynolds_perceptions_2014, the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007]. The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs). They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "... people are rewarded for their skills?".

Is the belief in meritocracy a perception or a preference with normative meaning? To expand the analytical conceptual framework, we believe that both dimensions should be considered, as proposed by @duru-bellat_who_2012. Such an approach opens up the possibility of analyzing whether perceptions and preferences are related (i.e., have a high correlation), or constitute independent aspects of the same phenomenon. As @sonhing_merit_2011 has pointed out, "People can believe that outcomes ought to be distributed based on merit and yet vary in their perceptions of whether this is how society currently operates" (p. 435). In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something different to someone perceiving high meritocracy than to someone perceiving low meritocracy. To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].

@@

Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, for instance, family status and the use of personal advantages (such as contacts or having a wealthy family) to get ahead in life. This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.  For instance, @kunovich_systems_2007 used an items' battery following the question “How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy). They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued as 1 when rated as "not important at all" (i.e. reverse coded). The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic aspect (such as having family and/or children) implies a stronger belief in meritocracy. A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].

The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above. They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections were classified as non-meritocratic. Despite making this distinction, in the operationalization the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did. Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items. Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment. For instance, a such distinction could open the door to studying to what extent different and contradictory views of meritocracy can coexist, to what extent individuals and societies differ in this regard, and of course the associated factors to such differences. Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of other terms like ascription, given that this last one leaves out aspects such as the use of personal relationships as a way to achieve success.

@@

Finally, most meritocracy studies have not adequately considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy. Such a strategy assumes that the latent construct is measured perfectly (i.e., no error or residual variance) by the selected indicators, going as far as to propose that "... In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654]. Although Reynolds and Xian [-@reynolds_perceptions_2014] made advances by conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.

There is a relevant trade-off regarding measurement error estimation in survey scales. Including multiple items per conceptual dimension opens the possibility of validity studies (at least in terms of the construct) by using factor analysis techniques, particularly confirmatory ones. However, increasing the number of items reduces the probability of usability of the scale in survey research given space limitations. This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted for a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.

## A conceptual proposal for studying meritocracy {-#instrumentprop}

@@

- Multiple indicators for each dimension to _account for measurement error_ in a confirmatory factor analysis context.

- Based on _previous indicators_ as much as possible to keep comparability between studies. As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we propose an operationalization suitable for this dataset, as well as a novel scale that attempts to rescue as much as possible the original wording of ISSP items.

- _Brief_, as to be used in regular public opinion surveys [@davidov_measurement_2009]. In this regard, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).

@@

```{r merit-model, echo=FALSE, fig.cap = "Conceptual Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
 knitr::include_graphics("input/images/Fig1.generalf.png")
```

The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019]. Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as "Hard work is important to get ahead in society". Preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research. Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to get ahead in life.

@@

The analysis is organized into three studies. Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy conceptual model with available secondary data from the inequality module of the International Social Survey Programme (ISSP). Study 2 consists of the proposal of a new scale for measuring meritocratic perceptions and preferences. Finally, in Study 3 we perform an additional validity analysis of the meritocracy scale.

# Study 1: Analyzing meritocracy with ISSP data

@@

The data corresponds to the social inequality module from the International Social Survey Programme (ISSP), the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues [  @issp-research-groupInternationalSocialSurvey2017]. This wave corresponds to 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals. Although there is data available from this module for a more recent wave (2019), the 2009 wave contains relevant items for studying meritocracy that unfortunately are not present in the last wave. Table \@ref(tab:countries) shows details about the sample of countries.

@@

There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy conceptual scheme, as presented below in Table \@ref(tab:table-issp):

```{r table-issp, echo=FALSE}
 # if (!require("pacman")) install.packages("pacman")
 # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
library(knitr)
library(kableExtra)
library(dplyr)
#  pacman::p_load(knitr, kableExtra, dplyr)
  table_format = if(knitr::is_html_output()) {
    "html"
  } else if(knitr::is_latex_output()) {
    "latex"
  }

@@

<!-- # Appendix {-} -->
